\chapter*{Abstract}
Chronic kidney disease (CKD) is a progressive renal disorder associated with nephron loss, inflammation, extracellular matrix deposition, and irreversible decline in kidney function. Acute kidney injury (AKI) may resolve after transient damage, but incomplete repair can increase the risk of CKD progression. This thesis presents a machine learning-based computational framework for identifying candidate protein biomarker genes associated with AKI, CKD, and CKD progression using single-cell RNA sequencing data from the GSE183276 Human Kidney Healthy and Injury Cell Atlas. Raw count data were processed using quality control filtering, library-size normalization, log transformation, and highly variable gene selection. The final feature matrix consisted of the top 3000 highly variable genes ranked by variance, with single cells as samples and \texttt{condition.l1} labels representing reference, AKI, and CKD conditions. Random Forest classifiers were trained using balanced class weights, and feature importance scores were combined with absolute expression differences to prioritize biomarkers. Pairwise analyses were performed for AKI versus reference, CKD versus reference, and CKD versus AKI. The results emphasize biologically meaningful injury and fibrosis-associated genes, including \textit{HAVCR1}, \textit{LCN2}, \textit{COL1A1}, \textit{FN1}, and \textit{SERPINE1}. The proposed workflow demonstrates how interpretable machine learning and single-cell transcriptomics can support biomarker discovery for kidney disease.

\textbf{Keywords:} chronic kidney disease, acute kidney injury, single-cell RNA sequencing, Random Forest, biomarker discovery, highly variable genes, kidney fibrosis.
\clearpage

\chapter*{Acknowledgement}
I express sincere gratitude to my supervisor for guidance, feedback, and encouragement throughout this thesis. I also acknowledge the researchers who generated and made publicly available the GSE183276 Human Kidney Healthy and Injury Cell Atlas dataset. Finally, I thank my department, teachers, family, and peers for their support during the completion of this work.
\clearpage
